---
title: "random_stuff"
author: "Logan Kelly, Ph.D."
date: "1/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# `jtools` `summ()` Will not work in data camp. add for shiny

Let's use a package called `jtools` to display the summary of the model. `jtools` has a function call `summ()` that is a bit nicer than `summary()`. We need to load the package using the `library()` function. I like to use the following code to load packages

```{r echo=T, eval=F}
if (!require("jtools")) install.packages("jtools")
  suppressMessages(library(jtools))
```

This first line tests id the package has been installed, if it has not, then R installs the package. the next line quietly loads the package. So the following code loads `jtools` and summarizes the model.

```{r, echo=T,eval=T}
if (!require("jtools")) install.packages("jtools")
  suppressMessages(library(jtools))
  
jtools::summ(mod1)
```

::: {.info-box .warning}
Note that the `summ()` function has two m's. This is to differentiated from a function that adds numbers up! 
:::